---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# checknormality

<!-- badges: start -->

[![Travis build
status](https://travis-ci.com/chrsshn/checknormality.svg?branch=main)](https://travis-ci.com/chrsshn/checknormality)
[![Codecov test
coverage](https://codecov.io/gh/chrsshn/checknormality/branch/main/graph/badge.svg)](https://codecov.io/gh/chrsshn/checknormality?branch=main)
<!-- badges: end -->

The goal of checknormality is to provide an implementation of popular normality tests that returns the test statistics from various normality tests. Comparable versions of the Shapiro-Wilk test and the Kolmogorov-Smirnov test are accessible through the base stats package (base::shapiro.test and base::ks.test, respectively). As of 2020/11/28, only the 

## A Note on the Algorithms

### Shapiro-Wilk Test
- For n > 50, the J. P. Royston approach for the Shapiro-Wilk test as described [here](https://www.real-statistics.com/tests-normality-and-symmetry/statistical-tests-normality-symmetry/shapiro-wilk-expanded-test/) must be used

- For  3 <= n <= 50, there are three "options": the original approach for the Shapiro-Wilk test as described [here](https://www.real-statistics.com/tests-normality-and-symmetry/statistical-tests-normality-symmetry/shapiro-wilk-test/), a modified approach that is compatible with the Royston approach as described in the last paragraph [here](https://www.real-statistics.com/tests-normality-and-symmetry/statistical-tests-normality-symmetry/shapiro-wilk-test/), and the Royston approach for n > 20

### Inverse of Normal CDF (Phi inverse)
- The Royston approach requires the inverse of the normal CDF, which is called in R with stats::qnorm(p)
- In the Rcpp implementation of the Royston approach, a [freely-available stand-alone C++ implementation](https://www.johndcook.com/blog/cpp_phi_inverse/) was used


## Installation

You can install the released version of checknormality from
[CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("checknormality")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("chrsshn/checknormality")
```

## Example 1: Testing a sample from a normal distribution using the Royston and modified approaches
```{r echo=TRUE, eval = T}
library (checknormality)
set1 <- rnorm (30, 0, 1)
shapiro.test (set1)
sw_test (set1, "royston")
sw_test (set1, "modified")
plot(density (set1))
```

## Example 2: Testing a sample from a non-normal distribution using the Royston and modified approaches
``` {r echo = T, eval = T}
library (checknormality)
set2 <- rpois(40, .787)
shapiro.test (set2)
sw_test (set2, "royston")
sw_test (set2, "modified")
plot(density (set2))
```

