---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
) 
```

# checknormality: because you should always check your assumptions

<!-- badges: start -->

[![Travis build
status](https://travis-ci.com/chrsshn/checknormality.svg?branch=main)](https://travis-ci.com/chrsshn/checknormality)
[![Codecov test
coverage](https://codecov.io/gh/chrsshn/checknormality/branch/main/graph/badge.svg)](https://codecov.io/gh/chrsshn/checknormality?branch=main)
<!-- badges: end -->

## Overview
The goal of checknormality is to provide implementations of popular normality tests that return the test statistics and p-values. As of 2020/11/30, the package contains an implementation of the [Shapiro-Wilk Test of Normality](https://en.wikipedia.org/wiki/Shapiro%E2%80%93Wilk_test).

## A Note on the Algorithms

### Shapiro-Wilk Test
- For n > 50, the J. P. Royston approach for the Shapiro-Wilk test as described [here](https://www.real-statistics.com/tests-normality-and-symmetry/statistical-tests-normality-symmetry/shapiro-wilk-expanded-test/) must be used.

- For 3 <= n <= 50, there are three "approaches": the original approach for the Shapiro-Wilk test as described [here](https://www.real-statistics.com/tests-normality-and-symmetry/statistical-tests-normality-symmetry/shapiro-wilk-test/), a modified approach that is compatible with the Royston approach as described in the last paragraph [here](https://www.real-statistics.com/tests-normality-and-symmetry/statistical-tests-normality-symmetry/shapiro-wilk-test/), and the Royston approach for n > 20.

### Inverse of Normal CDF (Phi inverse)
- The Royston approach requires the inverse of the normal CDF, which is calculated in R using
``` r
stats::qnorm(p)
```
- In the Rcpp implementation of the Royston approach, a [freely-available stand-alone C++ implementation](https://www.johndcook.com/blog/cpp_phi_inverse/) was used.


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("chrsshn/checknormality")
```

## Usage
### Example 1: Testing a small sample from an exponential distribution using the original, modified, and Royston approaches
```{r echo=TRUE, eval = T}
set1 <- rexp (30, .8)
checknormality::sw_test (set1, "original")
checknormality::sw_test (set1, "modified")
checknormality::sw_test (set1, "royston")
stats::shapiro.test (set1) 
plot(density (set1))
```

Note: you can also get the W test statistic using the functions checknormality::R_get_W () and checknormality::C_get_W() 

### Example 2: Testing a small sample from a normal distribution using the original, modified, and Royston approaches
``` {r echo = T, eval = T}
set2 <- rnorm (40, 20, 5)
checknormality::sw_test (set2, "original")
checknormality::sw_test (set2, "modified")
checknormality::sw_test (set2, "royston")
stats::shapiro.test (set2)
plot(density (set2))   
 
```

### Example 3: Testing a large sample from a Poisson distribution using the Royston approach
note: the original and modified approaches are only valid for n < 50 points
``` {r echo = T, eval = T}
set3 <- rpois(4000, .787)
checknormality::sw_test (set3, "royston")
stats::shapiro.test (set3)
plot(density (set3))   
 
```

## Example 4: Testing a large sample from a normal distribution using the Royston approach
note: the original and modified approaches are only valid for n < 50 points
``` {r echo = T, eval = T}
set4 <- rnorm(4000)
checknormality::sw_test (set4, "royston")
stats::shapiro.test (set4)
plot(density (set4))   
 
```

